A cautionary note on the use of Ornstein Uhlenbeck and other models in macroevolutionary studies

### 2015-04-17 Manuscript/ohyou.tex
Added ntaxa figure

% I ended up just LaTeXing this for our own amusement
      \begin{table}[!htbp]
        \begin{center}
        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model on a range of constant-rate birth death trees. Tree type refers to the extinction fraction for the birth-death trees. The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
        \bigskip
        \input{Tables/OU_bd_noerror.tex}
        \label{tables:bd_noerror}
        \end{center}
      \end{table}

@@

    We find that Type I error rates are unacceptably high when tree size is small (Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}), i.e., the OU model is often favoured, even though the Brownian model was actually used to generate the data. 
    For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains $>0.05$ even for trees with 1000 tips (Table \ref{tables:varrate_noerror}).
    This has clear implications for studies with low sample sizes where OU is favoured.

@@

\section{Bayesian methods (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)}

@@

\section{Just an OU problem? - Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}

>>>>

% I ended up just LaTeXing the tables for my own amusement
      \begin{table}[!htbp]
        \begin{center}
        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model on a range of constant-rate birth death trees. Tree type refers to the extinction fraction for the birth-death trees. The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
        \bigskip
        \input{Tables/OU_bd_noerror.tex}
        \label{tables:bd_noerror}
        \end{center}
      \end{table}

@@

    We find that Type I error rates are unacceptably high when tree size is small (Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}), i.e., the OU model is often favoured, even though the Brownian model was used to generate the data. 
    For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains $>0.05$ even for trees with 1000 tips (Table \ref{tables:varrate_noerror}).
    This shows that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases. 
    Unfortunately, OU models are often fitted to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58; Figure ; see Supplemental Material). % link to supp mat

    \begin{figure}[h]
      \centering
      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/NTaxa.png}
      \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014. Two studies with \morethan 3000 taxa have been omitted for clarity. See Supplementary Material for details %linkto supp mat
      }
      \label{figure.ntaxa}
    \end{figure}

@@

\section{Bayesian methods}
 (available in BayesTraits, diversitree, others?; simulations back in; improves error rates but difficult challenge of choosing the right priors)

@@

\section{Just an OU problem?}
Other parameters have similar problems (reduced set of our simulations for kappa, delta, lambda, e.g. just with Yule trees an error)}
